---
title: "Group Assignment Skills: Data Prep.&Workflow Mgt"
author: "Team 5"
date: "2024-09-06"
output: pdf_document
---

## Team 5
#Sophie van Hest
#Eveline Cai
#Mette Swanenberg
#Tyamo van der Ceelen

## Research Motivation

Casting A-list actors is a strategic move to boost a film's visibility, financial success, and box office performance. Studies show that casting major stars, particularly Oscar winners, can significantly raise audience interest and market expectations, often increasing predicted box office revenues by millions. These actors not only influence box office results but also impact IMDb ratings, which shape consumer behavior, especially on streaming platforms. The research question, **"Does an actor influence IMDb popularity ratings of a movie?"**, can guide production and marketing strategies, underscoring the importance of star power in driving both film success and audience decisions.

## Data
Data1 includes:

- nconst (string) - alphanumeric unique identifier of the name/person
- primaryName (string)– name by which the person is most often credited
- birthYear – in YYYY format
- deathYear – in YYYY format if applicable, else '\\N'
- primaryProfession (array of strings)– the top-3 professions of the person
- knownForTitles (array of tconsts) – titles the person is known for

Data2 includes:

- tconst (string) - alphanumeric unique identifier of the title
- averageRating – weighted average of all the individual user ratings
- numVotes - number of votes the title has received

```{r echo=FALSE, warning=FALSE}
url1 <- "https://datasets.imdbws.com/name.basics.tsv.gz"
url2 <- "https://datasets.imdbws.com/title.ratings.tsv.gz"

file1 <- "name.basics.tsv.gz"
file2 <- "title.ratings.tsv.gz"

if (!file.exists(file1)) {
  download.file(url1, file1)
}
if (!file.exists(file2)) {
  download.file(url2, file2)
}

data1 <- read.csv(gzfile(file1), sep = "\t")
data2 <- read.csv(gzfile(file2), sep = "\t")

head(data1)
head(data2)
```

```{r echo=FALSE, warning=FALSE}
str(data1)
str(data2)
```
```{r echo=FALSE, warning=FALSE}
summary(data1$birthYear)
```

```{r echo=FALSE, warning=FALSE}
summary(data2$averageRating)
```
```{r echo=FALSE, warning=FALSE}
summary(data2$numVotes)
```

## Data exploration

```{r echo=FALSE, warning=FALSE}
#Libraries:
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
```
Plot of the distribution of birth years
```{r echo=FALSE, warning=FALSE}
data1$birthYear <- as.numeric(data1$birthYear)
ggplot(data1, aes(x = birthYear)) +
  geom_histogram(binwidth = 10, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Birth Years", x = "Birth Year", y = "Count") +
  theme_minimal() +
  scale_x_continuous(limits = c(1750, NA))
```
Plots of the distribution of average movie ratings and number of votes
```{r echo=FALSE, warning=FALSE}
# Plotting distribution of average movie ratings
ggplot(data2, aes(x = averageRating)) +
  geom_histogram(binwidth = 0.5, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Average Movie Ratings", x = "Average Rating", y = "Count") +
  theme_minimal()

# Plotting distribution of number of votes
ggplot(data2, aes(x = numVotes)) +
  geom_histogram(binwidth = 0.1, fill = "lightpink", color = "black") +
  labs(title = "Distribution of Number of Votes", x = "Number of Votes", y = "Count") +
  theme_minimal() +
  scale_x_log10()
```














